Water desalination using polyelectrolyte hydrogel: Gibbs ensemble modeling

### 2022-09-05 gibbs.tex
Update gibbs.tex

\paragraph{Compression in open system}
\begin{figure*}[t]
\subfloat[The gel partial pressure vs gel molar volume \label{fig: PV}]
{\includegraphics[width=0.48\textwidth]{figures/fig_PV_pub.pdf}}
\hspace{0.02\textwidth}
\subfloat[Supernate salinity vs gel molar volume\label{fig: CV}]
{\includegraphics[width=0.48\textwidth]{figures/fig_CV_pub.pdf}}
\caption{
The compression of the gel in the \emph{open system} (solid lines) and in the \emph{closed system} (dotted lines). 
Each solid curve corresponds to different salinity of the reservoir $\cs$ (see legend).
The shadowed areas limit the states with applied pressure below zero and above 5 bar.\label{fig: PV and CV}}
\end{figure*}
First, we run a set of simulations modeling the gel compression in \emph{open system}, \ie in equilibrium with a big bath of certain salinity, $\cs$. 
The simulations are run for a set of different simulation box volumes. 
Each simulation returns the averaged pressure, $P$, and the number of $\cl$ ions, which are present in the simulation box, $\ncl$.

In order to get the partial pressure of the gel, we substitute from $P$ the osmotic pressure of ions in the reservoir: $\Pgel=P - \Pout$.
$\Pout$, in turn, we calculate by running a separate simulation of a reservoir, containing ionic gas in equilibrium with the bath of the same salinity $\cs$. 
The gel partial pressure, $\Pgel$, is the pressure that needs to be applied to the gel via a solvent permeable filter to compress the gel to a specific molar volume, $\Vgel$. 
By gel molar volume $\Vgel$ we assume the volume of the gel calculated per on mol of gel monomer units.

The dependencies of $\Pgel$ on $\Vgel$ are present in Figure \ref{fig: PV} for a set of different salinities. 
These dependencies (the case of \emph{open system}) are represented by solid lines. 
For example, the blue solid line illustrates the compression (or swelling) of the gel in equilibrium with a reservoir of salinity, $\cs=0.063$ mol/l. 
The points where the pressure equals zero, $\Pgel=0$, (indicated by filled circles) are the gel \emph{free swelling equilibrium} states. 
These states shift towards smaller volumes with increase of salinity. 
In general, the increase of salinity shift all the $\Pgel(V)$ curves towards smaller volumes.
This effect is well known and is typical for all branched \emph{strong} polyelectrolytes. 
It is caused by the decrease of ions osmotic pressure and by a screening of electrostatic interactions \cite{Zhulina2000, Landsgesel2020a}.
(The salinity dependence of the size of \emph{weak} polyelectrolyte gel is in general non-monotonic. We discuss this case in \cite{Rud2018}.)

>>>>

\paragraph{Compression in open system}
\begin{figure*}[t]
\subfloat[The gel partial pressure vs gel molar volume \label{fig: PV}]
{\includegraphics[width=0.48\textwidth]{figures/fig_PV_pub.pdf}}
\hspace{0.02\textwidth}
\subfloat[Supernate salinity vs gel molar volume\label{fig: CV}]
{\includegraphics[width=0.48\textwidth]{figures/fig_CV_pub.pdf}}
\caption{
The compression of the gel in the \emph{open system} (solid lines) and in the \emph{closed system} (dotted lines). 
Each solid curve corresponds to different salinity of the reservoir $\cs$ (see legend).
The shadowed areas limit the states with applied pressure below zero and above 5 bar.\label{fig: PV and CV}}
\end{figure*}
First, we ran a set of simulations modeling the gel compression in \emph{open system}, \ie in equilibrium with a big bath of certain salinity, $\cs$. 
The simulations were run for a set of different simulation box volumes. 
Each simulation provided pressure, $P$, and the number of $\cl$ ions, which are present in the simulation box, $\ncl$.

In order to get the partial pressure of the gel, we subtracted from $P$ the osmotic pressure of ions in the reservoir: $\Pgel=P - \Pout$.
$\Pout$, in turn, we obtained from a separate simulation of a reservoir, containing ionic gas in equilibrium with the bath of the same salinity $\cs$. 
The gel partial pressure, $\Pgel$, is the pressure that needs to be applied to the gel via a solvent permeable filter to compress the gel to a specific molar volume, $\Vgel$. $\Vgel$ is the gel volume per one mole of gel monomer units.

These dependencies of $\Pgel$ on $\Vgel$ for \emph{open system} for a set of various salinities are presented
in Figure \ref{fig: PV} by solid lines. 
For example, the blue solid line illustrates the compression (or swelling) of the gel in equilibrium with a reservoir of salinity, $\cs=0.063$ mol/l. 
The points where the pressure equals zero, $\Pgel=0$, (indicated by filled circles) are the gel \emph{free swelling equilibrium} states. 
These states shift towards smaller volumes with increase of salinity. 
In general, the increase of salinity shift all the $\Pgel(V)$ curves towards smaller volumes.
This effect is well known and is typical for all branched \emph{strong} polyelectrolytes. 
It is caused by the decrease of ions osmotic pressure and by a screening of electrostatic interactions \cite{Zhulina2000, Landsgesel2020a}.
(The salinity dependence of the size of \emph{weak} polyelectrolyte gel is in general non-monotonic. We discuss this case in \cite{Rud2018}.)
